---
title: "Phylogenetic tree of responding OTUs"
author: "Jared Flater"
date: "1/15/2020"
output:
  html_document:
    df_print: paged
    fig_width: 8.5
    fig_height: 11
    fig_caption: true
    toc: true
    toc_float: true
    number_sections: false
    code_folding: hide
---
```{r setup}
knitr::opts_knit$set(root.dir = normalizePath("."))
```

```{r, echo = F, message = F, warning = F, eval=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

# The following initializes usage of Bioc devel
BiocManager::install(version='devel')

BiocManager::install("ggtree")
```

```{r, echo = F, message = F, warning = F}
library(reshape2)
library(tidyverse)
library(phyloseq)
library(ggtree)
library(viridis)
library(kableExtra)
library(vegan)
library(venn)
```
# Introduction
We would like to visualize the phylogenetic relationship among the amendment responders. We do this by sub setting our data to those OTUs with a LFC of 4 in the early and late groups of each treatment. 

* A tree file is necessary for this
```{r, message=F, error=F, warning=F}
# We have a tree for this data
tree <- read.tree("../data/tree.nwk")
# phyloseq object
inc.physeq <- readRDS("../data/physeq2020.RDS")
# Combine the two
inc <- merge_phyloseq(inc.physeq, tree)
# Subset to desired samples
inc <- subset_samples(inc, day %in% c("7", "14", "21", "35", "49", "97")) %>%
  filter_taxa(function(x) sum(x) >= 1, T) 

inc10 <- inc %>%
  filter_taxa(function(x) sum(x) >= 10, T)
#Below lines for posterity, I forget how to drop taxa sometimes
#no.unclass <- subset_taxa(inc, !Phylum=="Bacteria_unclassified")
#no.unclass <- subset_taxa(no.unclass, !Genus=="Gp6_unclassified")

rm(tree, inc.physeq)
```
```{r}
inc
inc10
```
2/3s of taxa are observed less than 10 times in the incubated samples...is this meaningful? 

# Tree     

This tree is showing comparisons between amendment and reference, the previous example showed differences between amendment and reference plus amendment from the previous time point. This should be a more accurate representation of OTUs responding to the amendment for each 

```{r, message=F, error=F, warning=F}
# RDS saved from AmendmentDESeq.Rmd
resp_alf <- readRDS("../data/LFC_alf_OTUs.RDS") %>% dplyr::rename(label = OTU)
resp_comp <- readRDS("../data/LFC_comp_OTUs.RDS") %>% dplyr::rename(label = OTU)
resp_mix <- readRDS("../data/LFC_mix_OTUs.RDS") %>% dplyr::rename(label = OTU)

resp <- rbind(resp_alf[1], resp_comp[1], resp_mix[1])

dist_all_resp <- distinct(resp)

df <- resp_alf %>%
  full_join(resp_comp) %>%
  full_join(resp_mix)
```

```{r, message=F, error=F, warning=F}
inc_resp <- prune_taxa(c(df$label), inc) %>%
  filter_taxa(function(x) sum(x) >= 1, T)
```

```{r}
#### Alfalfa ###
alf_resp_early <- subset_samples(inc, Treatment_Response %in% "Alfalfa_early") %>%
  filter_taxa(function(x) sum(x) >= 1, T) %>%
  transform_sample_counts(function(x) {x/sum(x)}) %>%
  prune_taxa(c(resp_alf$label), .) %>%  
  psmelt() %>%
  group_by(OTU) %>%
  summarise(Alfalfa_early_Abundance = mean(Abundance)) 

alf_resp_late <- subset_samples(inc, Treatment_Response %in% "Alfalfa_late") %>%
  filter_taxa(function(x) sum(x) >= 1, T) %>%
  transform_sample_counts(function(x) {x/sum(x)}) %>%
  prune_taxa(c(resp_alf$label), .) %>% 
  psmelt() %>%
  group_by(OTU) %>%
  summarise(Alfalfa_late_Abundance = mean(Abundance)) 

#### Compost ###
comp_resp_early <- subset_samples(inc, Treatment_Response %in% "Compost_early") %>%
  filter_taxa(function(x) sum(x) >= 1, T) %>%
  transform_sample_counts(function(x) {x/sum(x)}) %>% 
  prune_taxa(c(resp_comp$label), .) %>% 
  psmelt() %>%
  group_by(OTU) %>%
  summarise(Compost_early_Abundance = mean(Abundance)) 

comp_resp_late <- subset_samples(inc, Treatment_Response %in% "Compost_late") %>%
  filter_taxa(function(x) sum(x) >= 1, T) %>%
  transform_sample_counts(function(x) {x/sum(x)}) %>%
  prune_taxa(c(resp_comp$label), .) %>% 
  psmelt() %>%
  group_by(OTU) %>%
  summarise(Compost_late_Abundance = mean(Abundance)) 

#### Mix ###
mix_resp_early <- subset_samples(inc, Treatment_Response %in% "Mix_early") %>%
  filter_taxa(function(x) sum(x) >= 1, T) %>%
  transform_sample_counts(function(x) {x/sum(x)}) %>% 
  prune_taxa(c(resp_mix$label), .) %>% 
  psmelt() %>%
  group_by(OTU) %>%
  summarise(Mix_early_Abundance = mean(Abundance)) 

mix_resp_late <- subset_samples(inc, Treatment_Response %in% "Mix_late") %>%
  filter_taxa(function(x) sum(x) >= 1, T) %>%
  transform_sample_counts(function(x) {x/sum(x)}) %>%
  prune_taxa(c(resp_mix$label), .) %>% 
  psmelt() %>%
  group_by(OTU) %>%
  summarise(Mix_late_Abundance = mean(Abundance)) 
```
```{r}
alf_abund <- alf_resp_early %>%
  left_join(alf_resp_late, by = "OTU")
alf_abund

comp_abund <- comp_resp_early %>%
  left_join(comp_resp_late, by = "OTU")
comp_abund

mix_abund <- mix_resp_early %>%
  left_join(mix_resp_late, by = "OTU")
mix_abund

# abund <- rbind(alf_abund[1], comp_abund[1], mix_abund[1])

abund <- alf_abund %>%
  full_join(comp_abund) %>%
  full_join(mix_abund) 

abund 

testtest <- df %>%
  dplyr::rename(OTU = label) %>%
  left_join(abund, by = "OTU")
write_csv(testtest, file = "../data/lfcandabund.csv")
```

```{r, message=F, error=F, warning=F}
#inc_resp <- phy_tree(inc_resp)
#inc_resp
tree <- ggtree(inc_resp)
tree
```
```{r, message=F, error=F, warning=F}
df2 = dplyr::mutate(testtest, newlab = paste(OTU, Phylum, Genus, sep='  '))
head(df2)
df5 <- mutate_if(df2, 
                is.character, 
                str_replace_all, pattern = "unclassified", replacement = "uc")
df6 <- mutate_if(df5, 
                is.character, 
                str_replace_all, pattern = "candidate_division", replacement = "C.D.") 
```
```{r, message=F, error=F, warning=F}
resp_p <- tree %<+% df6 
#resp_p  
test <- ggtree(resp_p$data, ) +
  geom_tiplab(aes(label = newlab), size=2, align=TRUE) 

test

```
```{r, message=F, error=F, warning=F}

colnames(df2)
head(df2$OTU)

df3 <- df2 %>%
  select(OTU, Alfalfa_Early = Alfalfa_early_log2FoldChange, Alfalfa_Late = Alfalfa_late_log2FoldChange, Compost_Early = Compost_early_log2FoldChange, Compost_Late = Compost_late_log2FoldChange, Mix_Early = Mix_early_log2FoldChange, Mix_Late = Mix_late_log2FoldChange) %>%
  column_to_rownames(var = "OTU")

df5 <- df2 %>%
  select(OTU, Alfalfa_Early = Alfalfa_early_Abundance, Alfalfa_Late = Alfalfa_late_Abundance, Compost_Early = Compost_early_Abundance, Compost_Late = Compost_late_Abundance, Mix_Early = Mix_early_Abundance, Mix_Late = Mix_late_Abundance) %>%
  column_to_rownames(var = "OTU")

colnames(df2)

df4 <- df2 %>%
  select(OTU, Phylum, Class, Order, Family, Genus, Alfalfa_Early = Alfalfa_early_log2FoldChange, Alfalfa_Late = Alfalfa_late_log2FoldChange, Compost_Early = Compost_early_log2FoldChange, Compost_Late = Compost_late_log2FoldChange, Mix_Early = Mix_early_log2FoldChange, Mix_Late = Mix_late_log2FoldChange) %>%
  column_to_rownames(var = "OTU")

write.csv(df3, file ="../data/resp_table.csv")
write.csv(df4, file ="../data/Table_5.csv")

final_resp <- test %>% 
  gheatmap(df3, offset= 7.4, width= 1.5, font.size = 2, colnames=T, colnames_angle=-45, hjust=0, color = "white", colnames_position = 'bottom') +
    scale_fill_viridis_c(option="D", name = "Log2FC")
final_resp
#install.packages("ggnewscale")
library(ggnewscale)
p2 <- final_resp + new_scale_fill()
final_resp <- gheatmap(p2, df5, offset= 9.5, width= 1.5, font.size = 2, colnames=T, colnames_angle=-45, hjust=0, color = "white") +
    scale_fill_viridis_c(option="A", name = "Relative abundance") 
final_resp
```

## Table of Otus    
All Otus with LFC greater than 4 in all treatment response groups
```{r, message=F, error=F, warning=F}
# Common responders in all amendments and response groups
new_DF <- df3[rowSums(is.na(df3)) == 0,]
kable(new_DF) %>%
  kable_styling(bootstrap_options = "striped") %>%
  scroll_box(width = "100%")
```

```{r}
tree_plot2 <- final_resp +
  geom_label2(aes(subset = !isTip, label=node), size=2, color="darkblue", alpha=1) 
ggsave("../Figures/resp_tree_allnodes.png", plot = tree_plot2, device = "png", width = 18.5, height = 32, units = "cm", dpi = 350)
tree_plot2
```
```{r}
# These nodes were interesting
new.labels <- data.frame(node = c(232,201,212,183,186,193,158,132,124,126), 
                         new.label = c(1,2,3,4,5,6,7,8,9,10))

p <- final_resp  + 
  geom_hilight(node = 232, fill = "gray", alpha=0.5, extend = 7.95) + 
  geom_hilight(node = 201, fill = "gray", alpha=0.5, extend = 8.09) + 
  geom_hilight(node = 212, fill = "gray", alpha=0.5, extend = 8.48) + 
  geom_hilight(node = 183, fill = "gray", alpha=0.5, extend = 8.27) + 
  geom_hilight(node = 186, fill = "gray", alpha=0.5, extend = 8.25) + 
  geom_hilight(node = 193, fill = "gray", alpha=0.5, extend = 8.3) + 
  geom_hilight(node = 158, fill = "gray", alpha=0.5, extend = 8.35) + 
  geom_hilight(node = 132, fill = "gray", alpha=0.5, extend = 8.59) + 
  geom_hilight(node = 124, fill = "gray", alpha=0.5, extend = 8.55) + 
  geom_hilight(node = 126, fill = "gray", alpha=0.5, extend = 8.57) 

pp <- p %<+% new.labels +
  geom_nodelab(aes(label=new.label), size=3, color="red", alpha = 1, nudge_x = .7) 
pp
ggsave("../Figures/resp_tree_wnodes.png", plot = pp, device = "png", width = 19, height = 32, units = "cm", dpi = 350)
```

# Responders orginating from amendment? 

Aliens??

```{r}
df3 <- read_csv("../data/resp_table.csv")
df3
treeOTUs <- df3$X1
alfALIENs <- readRDS("../data/alf.aliens.rds")
compALIENs <- readRDS("../data/comp.aliens.rds")
mixALIENs <- readRDS("../data/mix.aliens.rds")

venn(list(treeOTUs, alfALIENs, compALIENs, mixALIENs), snames = c("Responder", "Alf Aliens", "Compost Aliens", "Mix Aliens"))

venn(list(treeOTUs, alfALIENs), snames = c("Responder", "Alf Aliens"))
a <- venn(list(treeOTUs, alfALIENs), snames = c("Responder", "Alf Aliens"))


venn(list(treeOTUs, compALIENs), snames = c("Responder", "Compost Aliens"))
c <- venn(list(treeOTUs, compALIENs), snames = c("Responder", "Compost Aliens"))
```
```{r}
# Alfalfa
print("Alfalfa Aliens")
attr(a, "intersections")$`Responder:Alf Aliens`

# Compost
print("Compost Aliens")
attr(c, "intersections")$`Responder:Compost Aliens`
```
    
We have 3 OTUs from compost and alfalfa that are detected as responders! Let's label them on the tree plot

```{r}
source("functions.R")
# Make these green, orgianlly from alfalfa
p$data %>%
  filter(label %in% c("Otu01028", "Otu01189",  "Otu03173")) %>%
  select(node)
# Make these brown, orgianlly from compost
p$data %>%
  filter(label %in% c("Otu00887", "Otu00985", "Otu01490")) %>%
  select(node)

t <- pp + 
  geom_hilight(node = 4, fill = "green4", alpha= .5, extend = 1.07) + 
  geom_hilight(node = 24, fill = "green4", alpha= .5, extend = .73) + 
  geom_hilight(node = 32, fill = "green4", alpha= .5, extend = .75) + 
  geom_hilight(node = 62, fill = "brown", alpha= .5, extend = .7) + 
  geom_hilight(node = 64, fill = "brown", alpha= .5, extend = .745) + 
  geom_hilight(node = 65, fill = "brown", alpha= .5, extend = .79) + 
  scale_y_continuous(expand = expansion(mult = c(.05, 0))) +
  theme(legend.title = element_text(size = 6),legend.text = element_text(size = 6))
ggsave("../Figures/Figure_7.tiff", plot = t , device = "tiff", width = 140, height = 240, units = "mm", dpi = 500)

```

