---
title: "Soil Microcosm bacterial and chemical analysis"
output:
  html_document:
    df_print: paged
---
To understand the microbial response to amendments and to determine which microbes persist and are transfered from amendmen to soil, we need to characterize the microbial community in the amendments and starting soil. In addition, we will look at the inorganic N concentration in the inputs as well as the C:N ratio.
```{r echo=FALSE}
library(ggplot2)
library(vegan)
library(dplyr)
library(scales)
library(grid)
library(reshape2)
library(phyloseq)
library(tidyverse)
library(DESeq2)
library(gplots)
library(ape)
library(nlme)
library(emmeans)
library(xtable)
library(ggpubr)
library(purrr)
library(broom)
library(gridExtra)
```

We will subset to the samples of interest:
```{r eval=FALSE}
getwd()
#Read in the phyloseq object, raw object from mothur, see code file mothur_to_phyloseq.R
inc.raw.physeq <- readRDS("Data/incubation_physeq_Aug18.RDS")

#sample_data(inc.raw.physeq)

# We will be characterizing the ammendments and starting soil, exclude all else
inc.inputs <- subset_samples(inc.raw.physeq, treatment %in% c("AlfalfaAmend", "AlfalfaSoil", "CompostAmend"))
```
And take a look at the library size (# of OTUs) for each sample    
```{r}
df <- as.data.frame(sample_data(inc.inputs)) # Put sample_data into a ggplot-friendly data.frame
df$LibrarySize <- sample_sums(inc.inputs)
df <- df[order(df$LibrarySize),]
df$Index <- seq(nrow(df))
ggplot(data=df, aes(x=Index, y=LibrarySize, color=treatment)) + geom_point()
```
We need to add data for inorganic nitrogen concentration and C:N ratio. Inorganic N can be calculated by the sumation of `NH3` and `NO3` and C:N is `C_flash` / `N_flash`:
```{r}
# Pull out sample data and make a data.frame, pipe that into mutate() to create the new columns
df <- df %>%
  mutate(C_N = C_flash / N_flash, Inorganic_N = NH3 + NO3) 

rownames(df) <- df$i_id
sample_data(inc.inputs) <- df
# Sample names have start with a 2, we need to change these
names <- gsub(pattern = "2mm\\.", replacement = "", x = sample_names(inc.inputs))
sample_names(inc.inputs) <- names
sample_names(inc.inputs)
```    
```{r}
colnames(df)
df
td <- df %>%
  select(N_flash, C_flash, NH3, NO3, Inorganic_N, C_N) 
td

d <- td[c("2mm.dry.alfalfa.soil", "2mm.dry.compost7", "2mm.dry.alfalfa.plant3"),] 
d
row.names(d) <- c("Soil", "Compost", "Alfalfa")

td.table <- xtable(d)
td.table
```
Let's check some of the of taxa counts from theses samples
```{r}
# Minimum sample size
min(taxa_sums(otu_table(inc.inputs)))

# Remove singletons and 0 counts
inc.inputs <- filter_taxa(inc.inputs, function(x) sum(x) >= 2, T)
min(taxa_sums(otu_table(inc.inputs)))
```

```{r}
inc.inputs 
physeq.dist <- vegdist(t(data.frame(otu_table(inc.inputs))), method = "bray")

PCoA_ord <- ordinate(
  physeq = inc.inputs, 
  method = "PCoA",
  distance = physeq.dist
)

input.PCoA <- plot_ordination(physeq = inc.inputs, ordination = PCoA_ord, color = "treatment", title = "Input PCoA") + 
  geom_point(size = 1.5) +
  theme_bw() +
  theme(legend.title = element_blank())

input.PCoA 
```

# Relative Abundance in inputs
```{r}
# Put phyloseq object into a df with .02% phylum (glomed at phylum level)
RelativeAbundanceDf <- function(physeq) {
    physeq %>% tax_glom(taxrank = "Phylum") %>% transform_sample_counts(function(x) {
        x/sum(x)
    }) %>% psmelt() %>% filter(Abundance > 0.02) %>% arrange(Phylum)
}

# Function to plot relative abundance
PlotRelativeAbundance <- function(df) {
    ggplot(df, aes(x = as.factor(treatment), y = Abundance, fill = Phylum)) + 
    geom_bar(stat = "identity") +
    theme(axis.title.x = element_blank()) + 
    guides(fill = guide_legend(reverse = TRUE, keywidth = 1, keyheight = 1)) + 
    ylab("Relative Abundance (Phyla > 2%) \n") +
    ggtitle("Phylum Composition of Incubation Microcosms") 
}

inc.merged <- merge_samples(inc.inputs, "treatment")
sample_data(inc.merged)$treatment <- levels(sample_data(inc.inputs)$treatment)

plot <- PlotRelativeAbundance(RelativeAbundanceDf(inc.merged))
plot

sample_data(rare6k.physeq)
rare.merged <- merge_samples(rare6k.physeq, "Treatment_Day")
sample_data(rare.merged)$Treatment_Day <- levels(sample_data(rare6k.physeq)$Treatment_Day)
relllll <- PlotRelativeAbundance(RelativeAbundanceDf(rare.merged)) +
  facet_grid(~ day, labeller = labeller(day = as_labeller(day_names))) +
  scale_x_discrete(labels = treatment_names) +
  rotate_x_text(angle = 45)

tiff("relllll.tif",height=4,width=6,units='in',res=1200)
relllll
dev.off()
treatment_names <- c(
                    `1` = "Alfalfa",
                    `2` = "Mix",
                    `3` = "Compost",
                    `4` = "Reference"
                    )
day_names <- c(
                    `1` = "0",
                    `2` = "7",
                    `3` = "14",
                    `4` = "21",
                    `5` = "35",
                    `6` = "49",
                    `7` = "97"
                    )
sample_data(rare.merged)
```
```{r}
df <- sample_data(inc.merged)
ggplot(df, aes(x= treatment, y = Inorganic_N)) +
  geom_bar(aes(color=treatment, fill=treatment), stat = "identity", position = "stack") +
  ggtitle("Inorganic Nitrogen concentrations in inputs")

ggplot(df, aes(x= treatment, y = C_N)) +
  geom_bar(aes(color=treatment, fill=treatment), stat = "identity", position = "stack") +
  ggtitle("Carbon to Nitrogen ratio in inputs")
```
```{r}
#Subset to only days of the incubation, leaving out raw soil and amendments
inc.physeq <- subset_samples(inc.raw.physeq, day %in% c("0", "7", "14", "21", "35", "49", "97"))
#Rename treatments to more informative titles
data <- data.frame(sample_data(inc.physeq)) %>%
  mutate(treatment = recode(treatment, 'Control' = 'Reference', 'CompAlfa' = 'Mix'))
rownames(data) <- data$i_id
sample_data(inc.physeq) <- data
sample_data(inc.physeq)$day <- as.factor(sample_data(inc.physeq)$day)

sample_data(inc.physeq)
df <- as.data.frame(sample_data(inc.physeq)) # Put sample_data into a ggplot-friendly data.frame
df$LibrarySize <- sample_sums(inc.physeq)
df <- df[order(df$LibrarySize),]
df$Index <- seq(nrow(df))
ggplot(data=df, aes(x=i_id, y=LibrarySize, color=treatment)) + geom_point() +
  facet_grid(treatment ~ day) +
  ggtitle("Lbrary size of sample by treatment and day, large variation") +
  geom_hline(yintercept = 6000)
```

# Rarefy to 6k based on rarefactioncurve.R
```{r}
rare6k.physeq <- rarefy_even_depth(inc.physeq, sample.size = 6000, rngseed = 15879966) %>%
  filter_taxa(function(x) sum(x) >= 2, T)
```
# Ordination
```{r}
####PCoA
#######################
PCOA <- function(physeq){
  physeq.dist <- vegdist(t(data.frame(otu_table(physeq))), method = "bray")
  PCoA_ord <- ordinate(
    physeq = physeq, 
    method = "PCoA",
    distance = physeq.dist
  )
  trt.PCoA <- plot_ordination(physeq = physeq, ordination = PCoA_ord, 
                                                   title = "Treatment PCoA") + 
    geom_point(color = "grey90", size = .25) +
    theme_bw() +
    theme(legend.title = element_blank())
  day.PCoA <- plot_ordination(physeq = physeq, ordination = PCoA_ord, 
                              title = "Day PCoA") + 
    geom_point(color = "grey90", size = .25) +
    theme_bw() +
    theme(legend.title = element_blank())
  ord.plot <- grid.arrange(
    trt.PCoA + stat_ellipse(geom = "polygon", type = "norm", alpha = 0.3, aes(fill = treatment)), 
    day.PCoA + stat_ellipse(geom = "polygon", type = "norm", alpha = 0.3, aes(fill = day)),
    nrow = 1
  )
}
PCOA(rare6k.physeq)

tmp.ps <- subset_samples(rare6k.physeq, day %in% c("7", "14", "21", "35", "49", "97")) %>%
  filter_taxa(function(x) sum(x) >= 2, T)
ord.phy <- ordinate(tmp.ps, method = "PCoA",
                    distance = vegdist(t(data.frame(otu_table(tmp.ps))), method = "bray"))
ord <- plot_ordination(tmp.ps, ord.phy, type = "samples", color = "treatment")
ord + stat_ellipse(geom = "polygon", type = "norm", alpha = 0.3, aes(fill = treatment)) + ggtitle("PCoA of the incubation microcosms grouped by treatment")

ord <- plot_ordination(tmp.ps, ord.phy, type = "samples", color = "day")
ord + stat_ellipse(geom = "polygon", type = "norm", alpha = 0.3, aes(fill = day)) + ggtitle("PCoA of the incubation microcosms grouped by day")
 
```
# Constrained Ordinations
```{r}
# There aint no stinking soil data on day 0, drop it
rare6k.physeq <- subset_samples(rare6k.physeq, day %in% c("7", "14", "21", "35", "49", "97")) %>%
  filter_taxa(function(x) sum(x) >= 2, T)
# Need to add C:N and Inorganic_N to data for the rare object 
rd <- sample_data(rare6k.physeq)
rd <- rd %>%
  mutate(C_N = C_flash / N_flash, Inorganic_N = NH3 + NO3) 

rownames(rd) <- rd$i_id
sample_data(rare6k.physeq) <- rd

########################
# CAP ordinate
# All treatments
physeq.dist <- vegdist(t(data.frame(otu_table(rare6k.physeq))), method = "bray")
cap_ord <- ordinate(
  physeq = rare6k.physeq, 
  method = "CAP",
  distance = physeq.dist,
  formula = ~ pH + N_flash + C_flash + gravimetric_water_content + NH3 + NO3 + MBC_mg.kg_per_dry_wt_soil + C_N + Inorganic_N
)


# CAP plot
cap_plot <- plot_ordination(
  physeq = rare6k.physeq, 
  ordination = cap_ord, 
  color = "day", 
  axes = c(1,2)
) + 
  aes(shape = treatment) + 
  geom_point(aes(colour = day), alpha = 0.8, size = 4) + 
  geom_point(colour = "grey90", size = 1.5) 

# Now add the environmental variables as arrows
arrowmat <- vegan::scores(cap_ord, display = "bp")

# Add labels, make a data.frame
arrowdf <- data.frame(labels = rownames(arrowmat), arrowmat)

# Define the arrow aesthetic mapping
arrow_map <- aes(xend = CAP1, 
                 yend = CAP2, 
                 x = 0, 
                 y = 0, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

label_map <- aes(x = .8 * CAP1, 
                 y = .8 * CAP2, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

arrowhead = arrow(length = unit(0.02, "npc"))

# Make a new graphic
cap_plot + 
  geom_segment(
    mapping = arrow_map, 
    size = .5, 
    data = arrowdf, 
    color = "gray", 
    arrow = arrowhead
  ) + 
  geom_text(
    mapping = label_map, 
    size = 4,  
    data = arrowdf, 
    show.legend = FALSE
  ) + facet_grid(treatment ~ .)

########################
#### Alfalfa
inc <- subset_samples(rare6k.physeq, treatment %in% c("Alfalfa"))
inc
min(taxa_sums(inc))
inc <- filter_taxa(inc, function(x) sum(x) >= 2, T)

inc.dist <- vegdist(t(data.frame(otu_table(inc))), method = "bray")
cap_ord <- ordinate(
  physeq = inc, 
  method = "CAP",
  distance = inc.dist,
  formula = ~ pH + MBC_mg.kg_per_dry_wt_soil + C_N + Inorganic_N
)

# CAP plot
cap_plot <- plot_ordination(
  physeq = inc, 
  ordination = cap_ord, 
  color = "day", 
  axes = c(1,2)
) + 
  geom_point(aes(colour = day), alpha = 0.8, size = 4) + 
  geom_point(colour = "grey90", size = 1.5) 

# Now add the environmental variables as arrows
arrowmat <- vegan::scores(cap_ord, display = "bp")

# Add labels, make a data.frame
arrowdf <- data.frame(labels = rownames(arrowmat), arrowmat)

# Define the arrow aesthetic mapping
arrow_map <- aes(xend = CAP1, 
                 yend = CAP2, 
                 x = 0, 
                 y = 0, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

label_map <- aes(x = .8 * CAP1, 
                 y = .8 * CAP2, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

arrowhead = arrow(length = unit(0.02, "npc"))

# Make a new graphic
Alfalfa <- cap_plot + 
  geom_segment(
    mapping = arrow_map, 
    size = .5, 
    data = arrowdf, 
    color = "gray", 
    arrow = arrowhead
  ) + 
  geom_text(
    mapping = label_map, 
    size = 4,  
    data = arrowdf, 
    show.legend = FALSE
  ) +
  ggtitle("Alfalfa CAP")


########################
#### Compost
inc <- subset_samples(rare6k.physeq, treatment %in% c("Compost"))
inc
min(taxa_sums(inc))
inc <- filter_taxa(inc, function(x) sum(x) >= 2, T)

inc.dist <- vegdist(t(data.frame(otu_table(inc))), method = "bray")
cap_ord <- ordinate(
  physeq = inc, 
  method = "CAP",
  distance = inc.dist,
  formula = ~ pH + MBC_mg.kg_per_dry_wt_soil + C_N + Inorganic_N
)

# CAP plot
cap_plot <- plot_ordination(
  physeq = inc, 
  ordination = cap_ord, 
  color = "day", 
  axes = c(1,2)
) + 
  geom_point(aes(colour = day), alpha = 0.8, size = 4) + 
  geom_point(colour = "grey90", size = 1.5) 

# Now add the environmental variables as arrows
arrowmat <- vegan::scores(cap_ord, display = "bp")

# Add labels, make a data.frame
arrowdf <- data.frame(labels = rownames(arrowmat), arrowmat)

# Define the arrow aesthetic mapping
arrow_map <- aes(xend = CAP1, 
                 yend = CAP2, 
                 x = 0, 
                 y = 0, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

label_map <- aes(x = .8 * CAP1, 
                 y = .8 * CAP2, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

arrowhead = arrow(length = unit(0.02, "npc"))

# Make a new graphic
Compost <- cap_plot + 
  geom_segment(
    mapping = arrow_map, 
    size = .5, 
    data = arrowdf, 
    color = "gray", 
    arrow = arrowhead
  ) + 
  geom_text(
    mapping = label_map, 
    size = 4,  
    data = arrowdf, 
    show.legend = FALSE
  ) +
  ggtitle("Compost CAP")

########################
#### Mix
inc <- subset_samples(rare6k.physeq, treatment %in% c("Mix"))
inc
min(taxa_sums(inc))
inc <- filter_taxa(inc, function(x) sum(x) >= 2, T)

inc.dist <- vegdist(t(data.frame(otu_table(inc))), method = "bray")
cap_ord <- ordinate(
  physeq = inc, 
  method = "CAP",
  distance = inc.dist,
  formula = ~ pH + MBC_mg.kg_per_dry_wt_soil + C_N + Inorganic_N
)

# CAP plot
cap_plot <- plot_ordination(
  physeq = inc, 
  ordination = cap_ord, 
  color = "day", 
  axes = c(1,2)
) + 
  geom_point(aes(colour = day), alpha = 0.8, size = 4) + 
  geom_point(colour = "grey90", size = 1.5) 

# Now add the environmental variables as arrows
arrowmat <- vegan::scores(cap_ord, display = "bp")

# Add labels, make a data.frame
arrowdf <- data.frame(labels = rownames(arrowmat), arrowmat)

# Define the arrow aesthetic mapping
arrow_map <- aes(xend = CAP1, 
                 yend = CAP2, 
                 x = 0, 
                 y = 0, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

label_map <- aes(x = .8 * CAP1, 
                 y = .8 * CAP2, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

arrowhead = arrow(length = unit(0.02, "npc"))

# Make a new graphic
Mix <- cap_plot + 
  geom_segment(
    mapping = arrow_map, 
    size = .5, 
    data = arrowdf, 
    color = "gray", 
    arrow = arrowhead
  ) + 
  geom_text(
    mapping = label_map, 
    size = 4,  
    data = arrowdf, 
    show.legend = FALSE
  ) +
  ggtitle("Mix CAP")

########################
#### Control
inc <- subset_samples(rare6k.physeq, treatment %in% c("Reference"))
inc
min(taxa_sums(inc))
inc <- filter_taxa(inc, function(x) sum(x) >= 2, T)

inc.dist <- vegdist(t(data.frame(otu_table(inc))), method = "bray")
cap_ord <- ordinate(
  physeq = inc, 
  method = "CAP",
  distance = inc.dist,
  formula = ~ pH + MBC_mg.kg_per_dry_wt_soil + C_N + Inorganic_N
)

# CAP plot
cap_plot <- plot_ordination(
  physeq = inc, 
  ordination = cap_ord, 
  color = "day", 
  axes = c(1,2)
) + 
  geom_point(aes(colour = day), alpha = 0.8, size = 4) + 
  geom_point(colour = "grey90", size = 1.5) 

# Now add the environmental variables as arrows
arrowmat <- vegan::scores(cap_ord, display = "bp")

# Add labels, make a data.frame
arrowdf <- data.frame(labels = rownames(arrowmat), arrowmat)

# Define the arrow aesthetic mapping
arrow_map <- aes(xend = CAP1, 
                 yend = CAP2, 
                 x = 0, 
                 y = 0, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

label_map <- aes(x = .8 * CAP1, 
                 y = .8 * CAP2, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

arrowhead = arrow(length = unit(0.02, "npc"))

# Make a new graphic
Control <- cap_plot + 
  geom_segment(
    mapping = arrow_map, 
    size = .5, 
    data = arrowdf, 
    color = "gray", 
    arrow = arrowhead
  ) + 
  geom_text(
    mapping = label_map, 
    size = 4,  
    data = arrowdf, 
    show.legend = FALSE
  ) +
  ggtitle("Reference CAP")

grid.arrange(
  Control,
  Alfalfa,
  Compost,
  Mix,
  nrow = 2,
  top = "CAP Plots"
)
```

# Clustering Compost samples
```{r}
# Rarefy to 10k based on rarefactoncurve.R

compost.physeq <- subset_samples(rare6k.physeq, treatment %in% c("Compost")) %>%
  filter_taxa(function(x) sum(x) >= 2, T) 
# get day and i_id for these samples, we will use this list to order the days on the tree
compost.rownames <- data.frame(sample_data(compost.physeq)) %>%
  select(day, i_id)
# List of each samples day label to use as rownames for the clustering plot
rownamesforcompost <- as.character(compost.rownames$day)
# Distance matrix, setting binary to T compustes a presence absence for the OTU table
compost.physeq.dist <- vegdist(t(data.frame(otu_table(compost.physeq))), method = "bray", binary = T)
# Hierachical clustering to check if days are grouped
compost.clustering <- hclust(compost.physeq.dist, method = "complete")
compost.clustering$labels <- rownamesforcompost
plot(as.phylo(compost.clustering), type = "unrooted", use.edge.length = TRUE, col = "gray80", main = "Compost clustering analysis dendogram")
```
# Clustering Alfalfa samples
```{r}
alfalfa.physeq <- subset_samples(rare6k.physeq, treatment %in% c("Alfalfa")) %>%
  filter_taxa(function(x) sum(x) >= 2, T)
# get day and i_id for these samples, we will use this list to order the days on the tree
alfalfa.rownames <- data.frame(sample_data(alfalfa.physeq)) %>%
  select(day, i_id)
# List of each samples day label to use as rownames for the clustering plot
rownamesforalfalfa <- as.character(alfalfa.rownames$day)
# Distance matrix
alfalfa.physeq.dist <- vegdist(t(data.frame(otu_table(alfalfa.physeq))), method = "bray", binary = T)
# Hierachical clustering to check if days are grouped
alfalfa.clustering <- hclust(alfalfa.physeq.dist, method = "complete")
alfalfa.clustering$labels <- rownamesforalfalfa
plot(as.phylo(alfalfa.clustering), type = "unrooted", use.edge.length = TRUE, col = "gray80", main = "Alfalfa clustering analysis dendogram")

reference.physeq <- subset_samples(rare6k.physeq, treatment %in% c("Reference")) %>%
  filter_taxa(function(x) sum(x) >= 2, T) 
# get day and i_id for these samples, we will use this list to order the days on the tree
reference.rownames <- data.frame(sample_data(reference.physeq)) %>%
  select(day, i_id)
# List of each samples day label to use as rownames for the clustering plot
rownamesforreference <- as.character(reference.rownames$day)
# Distance matrix
reference.physeq.dist <- vegdist(t(data.frame(otu_table(reference.physeq))), method = "bray", binary = T)
# Hierachical clustering to check if days are grouped
reference.clustering <- hclust(reference.physeq.dist, method = "complete")
reference.clustering$labels <- rownamesforreference
plot(as.phylo(reference.clustering), type = "unrooted", use.edge.length = TRUE, col = "gray80", main = "Reference clustering analysis dendogram")

mix.physeq <- subset_samples(rare6k.physeq, treatment %in% c("Mix")) %>%
  filter_taxa(function(x) sum(x) >= 2, T)
# get day and i_id for these samples, we will use this list to order the days on the tree
mix.rownames <- data.frame(sample_data(mix.physeq)) %>%
  select(day, i_id)
# List of each samples day label to use as rownames for the clustering plot
rownamesformix <- as.character(mix.rownames$day)
# Distance matrix
mix.physeq.dist <- vegdist(t(data.frame(otu_table(mix.physeq))), method = "bray", binary = T)
# Hierachical clustering to check if days are grouped
mix.clustering <- hclust(mix.physeq.dist, method = "complete")
mix.clustering$labels <- rownamesformix
plot(as.phylo(mix.clustering), type = "unrooted", use.edge.length = TRUE, col = "gray80", main = "Mix clustering analysis dendogram")
```
 
We see three groupings for each treatment, these groups will be know as response gropus, baseline, early and late    
baseline = day 0
early = day 7, 14, 21
late = day 35, 49, 97
```{r}
# Use the un-standardized inc.physeq for this
# Add response group to meta
inc.physeq.data <- data.frame(sample_data(inc.physeq))
inc.physeq.data$response.group[inc.physeq.data$day == "0"] <- "baseline" 
inc.physeq.data$response.group[inc.physeq.data$day %in% c("7", "14", "21")] <- "early" 
inc.physeq.data$response.group[inc.physeq.data$day %in% c("35", "49", "97")] <- "late" 
inc.physeq.data <- inc.physeq.data %>%
  mutate(Treatment_Day = paste(treatment, day, sep = '_'))
rownames(inc.physeq.data) <- data$i_id
sample_data(inc.physeq) <- inc.physeq.data

rare6k.physeq.data <- data.frame(sample_data(rare6k.physeq))
rare6k.physeq.data$response.group[rare6k.physeq.data$day == "0"] <- "baseline" 
rare6k.physeq.data$response.group[rare6k.physeq.data$day %in% c("7", "14", "21")] <- "early" 
rare6k.physeq.data$response.group[rare6k.physeq.data$day %in% c("35", "49", "97")] <- "late" 
sample_data(rare6k.physeq) <- rare6k.physeq.data



ps_tmp <- get_top_taxa(alfalfa.physeq, n = 1000, relative = TRUE,
                       discard_other = FALSE, other_label = "Other")

ps_tmp <- name_taxa(ps_tmp, label = "Unkown", species = T, other_label = "Other")

fantaxtic_bar(ps_tmp, color_by = "Phylum", label_by = NULL,
              facet_by = "response.group", other_label = "Other") + theme(legend.position = "none")

```
# DESeq: Now we pass this to a DESeq2 function
```{r eval=FALSE}
# Functions for use with DESeq
who_diff_day <- function(DDS, choice1, choice2, phy.object){
  res = results(DDS, contrast = c("response.group", choice1, choice2), cooksCutoff = FALSE)
  #plotCounts(AlfalfaDDS, gene="OTU_311", intgroup="day")
  #Use above line to check if an OTU is increasing or decreasing depending on order of contrast
  alpha = 0.01
  #alpha = 0.1
  sigtab = res[which(res$padj < alpha), ]
  sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(phy.object)[rownames(sigtab), ], "matrix"))
  theme_set(theme_bw())
  scale_fill_discrete <- function(palname = "Set1", ...) {
    scale_fill_brewer(palette = palname, ...)
  }
  # Phylum order
  x = tapply(sigtab$log2FoldChange, sigtab$Phylum, function(x) max(x))
  x = sort(x, TRUE)
  sigtab$Phylum = factor(as.character(sigtab$Phylum), levels=names(x))
  # Genus order
  x = tapply(sigtab$log2FoldChange, sigtab$Genus, function(x) max(x))
  x = sort(x, TRUE)
  sigtab$Genus = factor(as.character(sigtab$Genus), levels=names(x))
  #ggplot(sigtab, aes(x=Genus, y=log2FoldChange, color=phylum)) + geom_point(size=2) + 
  #  theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=1.0)) +
  #  ggtitle("Day 0 to Day 7")
  return(sigtab)
}
# function plot log2FoldChange 
log_plot <- function(sigtab,t1){
  ggplot(sigtab, aes(x=Genus, y=log2FoldChange, color=Phylum)) + geom_point(size=2) + 
    coord_flip() +
    ggtitle(t1)
} 
```

```{r}
# Use inc.physeq, not rarefied as DESeq does this
alf.physeq <- subset_samples(inc.physeq, treatment %in% c("Alfalfa")) %>%
  filter_taxa(function(x) sum(x) >= 2, T)

log.plot.early.alf <- alf.physeq %>%
  phyloseq_to_deseq2( ~ response.group) %>%
  DESeq(test = "Wald", fitType = "local") %>%
  who_diff_day("early", "baseline", alf.physeq) %>%
  log_plot("Alfalfa OTUS in early group that are significantly changing compared to day 0")
log.plot.early.alf

log.plot.late.alf <- alf.physeq %>%
  phyloseq_to_deseq2( ~ response.group) %>%
  DESeq(test = "Wald", fitType = "local") %>%
  who_diff_day("late", "early", alf.physeq) %>%
  log_plot("Alfalfa OTUS in late group that are significantly changing compared to early group")
log.plot.late.alf
###
comp.physeq <- subset_samples(inc.physeq, treatment %in% c("Compost")) %>%
  filter_taxa(function(x) sum(x) >= 2, T)

log.plot.early.comp <- comp.physeq %>%
  phyloseq_to_deseq2( ~ response.group) %>%
  DESeq(test = "Wald", fitType = "local") %>%
  who_diff_day("early", "baseline", comp.physeq) %>%
  log_plot("Compost OTUS in early group that are significantly changing compared to day 0")
log.plot.early.comp

log.plot.late.comp <- comp.physeq %>%
  phyloseq_to_deseq2( ~ response.group) %>%
  DESeq(test = "Wald", fitType = "local") %>%
  who_diff_day("late", "early", comp.physeq) %>%
  log_plot("Compost OTUS in late group that are significantly changing compared to early group")
log.plot.late.comp
###
mix.physeq <- subset_samples(inc.physeq, treatment %in% c("Mix")) %>%
  filter_taxa(function(x) sum(x) >= 2, T)

log.plot.early.mix <- mix.physeq %>%
  phyloseq_to_deseq2( ~ response.group) %>%
  DESeq(test = "Wald", fitType = "local") %>%
  who_diff_day("early", "baseline", mix.physeq) %>%
  log_plot("Mix OTUS in early group that are significantly changing compared to day 0")
log.plot.early.mix

log.plot.late.mix <- mix.physeq %>%
  phyloseq_to_deseq2( ~ response.group) %>%
  DESeq(test = "Wald", fitType = "local") %>%
  who_diff_day("late", "early", mix.physeq) %>%
  log_plot("Mix OTUS in late group that are significantly changing compared to early group")
log.plot.late.mix
###
ref.physeq <- subset_samples(inc.physeq, treatment %in% c("Reference")) %>%
  filter_taxa(function(x) sum(x) >= 2, T)

log.plot.early.ref <- ref.physeq %>%
  phyloseq_to_deseq2( ~ response.group) %>%
  DESeq(test = "Wald", fitType = "local") %>%
  who_diff_day("early", "baseline", ref.physeq) %>%
  log_plot("Reference OTUS in early group that are significantly changing compared to day 0")
log.plot.early.ref

log.plot.late.ref <- ref.physeq %>%
  phyloseq_to_deseq2( ~ response.group) %>%
  DESeq(test = "Wald", fitType = "local") %>%
  who_diff_day("late", "early", ref.physeq) %>%
  log_plot("Reference OTUS in late group that are significantly changing compared to early group")
log.plot.late.ref
```
Let's grab the early samples and look at how the OTUs are distributed accross treatments with a venn diagram
```{r}
early.alf.otus <- log.plot.early.alf$data %>%
  rownames_to_column() %>%
  select(OTU = rowname, Phylum, Genus, log2FoldChange) %>%
  filter(log2FoldChange >= 2) %>%
  arrange(desc(log2FoldChange))
early.alf.otus 

###
early.comp.otus <- log.plot.early.comp$data %>%
  rownames_to_column() %>%
  select(OTU = rowname, Phylum, Genus, log2FoldChange) %>%
  filter(log2FoldChange >= 2)
early.comp.otus

###
early.mix.otus <- log.plot.early.mix$data %>%
  rownames_to_column() %>%
  select(OTU = rowname, Phylum, Genus, log2FoldChange) %>%
  filter(log2FoldChange >= 2)
early.mix.otus

###
early.ref.otus <- log.plot.early.ref$data %>%
  rownames_to_column() %>%
  select(OTU = rowname, Phylum, Genus, log2FoldChange) %>%
  filter(log2FoldChange >= 2)
early.ref.otus

venn <- venn(list(Alfalfa = early.alf.otus$OTU, Reference = early.ref.otus$OTU,
                  Mix = early.mix.otus$OTU, Compost = early.comp.otus$OTU))

```
Late responders:
```{r}
late.alf.otus <- log.plot.late.alf$data %>%
  rownames_to_column() %>%
  select(OTU = rowname, Phylum, Genus, log2FoldChange) %>%
  filter(log2FoldChange >= 2)
late.alf.otus 

###
late.comp.otus <- log.plot.late.comp$data %>%
  rownames_to_column() %>%
  select(OTU = rowname, Phylum, Genus, log2FoldChange) %>%
  filter(log2FoldChange >= 2)
late.comp.otus

###
late.mix.otus <- log.plot.late.mix$data %>%
  rownames_to_column() %>%
  select(OTU = rowname, Phylum, Genus, log2FoldChange) %>%
  filter(log2FoldChange >= 2)
late.mix.otus

###
late.ref.otus <- log.plot.late.ref$data %>%
  rownames_to_column() %>%
  select(OTU = rowname, Phylum, Genus, log2FoldChange) %>%
  filter(log2FoldChange >= 2)
late.ref.otus

venn2 <- venn(list(Alfalfa = late.alf.otus$OTU, Reference = late.ref.otus$OTU, Mix = late.mix.otus$OTU, Compost = late.comp.otus$OTU))
```

```{r}
alf.early.otus <- attr(venn, "intersections")$Alfalfa
early.alf <- subset_samples(rare6k.physeq, treatment %in% c("Alfalfa")) %>%
  subset_samples(day %in% c("0", "7", "14", "21")) 
only.early.alf <- subset_taxa(early.alf, rownames(tax_table(early.alf)) %in% alf.early.otus)

comp.early.otus <- attr(venn, "intersections")$Compost
early.comp <- subset_samples(rare6k.physeq, treatment %in% c("Compost")) %>%
  subset_samples(day %in% c("0", "7", "14", "21")) 
only.early.comp <- subset_taxa(early.comp, rownames(tax_table(early.comp)) %in% comp.early.otus)
```

```{r}
lotu <- paste(c(comp.early.otus, alf.early.otus))
compalf <- subset_samples(rare6k.physeq, treatment %in% c("Compost", "Alfalfa")) %>%
  subset_samples(day %in% c("0", "7", "14", "21")) 

only.early.compalf <- subset_taxa(compalf, rownames(tax_table(compalf)) %in% lotu) 

plot_bar(only.early.compalf, x = "response.group", fill = "Order", facet_grid = ~treatment) +
  geom_bar(aes(color=Order, fill=Order), stat = "identity", position = "stack")
```
Repeat for late vs early
```{r}
alf.late.otus <- attr(venn2, "intersections")$Alfalfa
late.alf <- subset_samples(rare6k.physeq, treatment %in% c("Alfalfa")) %>%
  subset_samples(day %in% c("7", "14", "21", "35", "49", "97")) 
only.late.alf <- subset_taxa(late.alf, rownames(tax_table(late.alf)) %in% alf.late.otus)

comp.late.otus <- attr(venn2, "intersections")$Compost
late.comp <- subset_samples(rare6k.physeq, treatment %in% c("Compost")) %>%
  subset_samples(day %in% c("7", "14", "21", "35", "49", "97")) 
only.late.comp <- subset_taxa(late.comp, rownames(tax_table(late.comp)) %in% comp.late.otus)
```
```{r}


rare6k.physeq <- transform_sample_counts(rare6k.physeq, function(x) {
        x/sum(x)})

lotu2 <- paste(c(comp.late.otus, alf.late.otus))
compalf.late <- subset_samples(rare6k.physeq, treatment %in% c("Compost", "Alfalfa")) %>%
  subset_samples(day %in% c("7", "14", "21", "35", "49", "97")) 

only.late.compalf <- subset_taxa(compalf.late, rownames(tax_table(compalf.late)) %in% lotu2)

library(fantaxtic)
# get top 10 taxa from the late responding group
late.compalfa.10 <- get_top_taxa(only.late.compalf, length(lotu2), relative = T, discard_other = F, other_label = "Other")
late.compalfa.10 <- name_taxa(late.compalfa.10, label = "Unknown", species = T, other_label = "Other")


fantaxtic_bar(only.late.compalf, color_by = "Genus", label_by = "Phylum", other_label = "Other", facet_by = "response.group", grid_by = "treatment")
sample_data(late.compalfa.10)

plot_bar(only.late.compalf, x = "response.group", fill = "Order", facet_grid = ~treatment) +
  geom_bar(aes(color=Order, fill=Order), stat = "identity", position = "stack")
```
]
```{r}

only.early.alf.data <- as.data.frame(tax_table(only.early.alf)) %>%
  rownames_to_column("OTU")
df <- merge(only.early.alf.data, early.alf.otus, by = "OTU") %>%
  select(OTU, Phylum.y, Genus.y, log2FoldChange) %>%
  arrange(desc(log2FoldChange))
df

write.table(df, file = "early.alf.resonders.txt", sep = ",", quote = FALSE, row.names = F)
pie(table(droplevels(df$Phylum.y)))


only.late.alf.data <- as.data.frame(tax_table(only.late.alf)) %>%
  rownames_to_column("OTU")
df <- merge(only.late.alf.data, late.alf.otus, by = "OTU") %>%
  select(OTU, Phylum.y, Genus.y, log2FoldChange) %>%
  arrange(desc(log2FoldChange))
df
write.table(df, file = "late.alf.resonders.txt", sep = ",", quote = FALSE, row.names = F)
pie(table(droplevels(df$Phylum.y)))

only.early.comp.data <- as.data.frame(tax_table(only.early.comp)) %>%
  rownames_to_column("OTU")
df <- merge(only.early.comp.data, early.comp.otus, by = "OTU") %>%
  select(OTU, Phylum.y, Genus.y, log2FoldChange) %>%
  arrange(desc(log2FoldChange))
df

write.table(df, file = "early.comp.resonders.txt", sep = ",", quote = FALSE, row.names = F)
pie(table(droplevels(df$Phylum.y)))

only.late.comp.data <- as.data.frame(tax_table(only.late.comp)) %>%
  rownames_to_column("OTU")
df <- merge(only.late.comp.data, late.comp.otus, by = "OTU") %>%
  select(OTU, Phylum.y, Genus.y, log2FoldChange) %>%
  arrange(desc(log2FoldChange))
df

write.table(df, file = "late.comp.resonders.txt", sep = ",", quote = FALSE, row.names = F)
# Paste this into the console 

tiff("late_comp_reponders_pie.tif",height=4,width=4,units='in',res=1200)
pie(table(droplevels(df$Phylum.y)))
dev.off()
```

```{r}

```

```{r}

```
# OTU Presence Calculation
```{r}
# Now calculate presence of each OTU
abund_val <- function(physeq){
  # Calculate the frequency of each OTU across all samples
  physeq = transform_sample_counts(physeq, function(x) {x[x >= (1)] <- 1 
  return(x)})
  otu.freq = rowSums(otu_table(physeq))/nsamples(physeq)
  phyla = as.vector(data.frame(tax_table(physeq)))
  otuabun = cbind.data.frame(frequency = otu.freq,phyla)
  return(otuabun)
}
early.comp.otu.freq <- abund_val(comp.taxa) %>%
  rownames_to_column() %>%
  filter(frequency >= .25)
early.comp.otu.freq  
```