---
title: "Phylogenetic tree of responding OTUs"
author: "Jared Flater"
date: "1/15/2020"
output:
  html_document:
    df_print: paged
    fig_width: 8.5
    fig_height: 11
    fig_caption: true
    toc: true
    toc_float: true
    number_sections: false
    code_folding: hide
---
```{r setup}
knitr::opts_knit$set(root.dir = normalizePath("."))
```

```{r, echo = F, message = F, warning = F, eval=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE))
  install.packages("BiocManager")

BiocManager::install("ggtree")
BiocManager::install("phyloseq")
BiocManager::install("treeio")
```

```{r, echo = F, message = F, warning = F}
library(reshape2)
library(tidyverse)
library(phyloseq)
library(ggtree)
library(treeio)
library(viridis)
library(kableExtra)
library(vegan)
library(venn)
```
# Introduction
We would like to visualize the phylogenetic relationship among the amendment responders. We do this by sub setting our data to those OTUs with a LFC of 4 in the early and late groups of each treatment. 

* A tree file is necessary for this
```{r, message=F, error=F, warning=F}
# We have a tree for this data
tree <- read.tree("../data/tree.nwk")
# phyloseq object
inc.physeq <- readRDS("../data/incubation_physeq_Aug18.RDS")
# Combine the two
inc <- merge_phyloseq(inc.physeq, tree)
# Subset to desired samples
inc <- subset_samples(inc, day %in% c("7", "14", "21", "35", "49", "97")) %>%
  filter_taxa(function(x) sum(x) >= 1, T) 

inc10 <- inc %>%
  filter_taxa(function(x) sum(x) >= 10, T)
#Below lines for posterity, I forget how to drop taxa sometimes
#no.unclass <- subset_taxa(inc, !Phylum=="Bacteria_unclassified")
#no.unclass <- subset_taxa(no.unclass, !Genus=="Gp6_unclassified")

rm(tree, inc.physeq)
```
```{r}
inc
inc10
```
2/3s of taxa are observed less than 10 times in the incubated samples...is this meaningful? 

# Tree     

This tree is showing comparisons between amendment and reference, the previous example showed differences between amendment and reference plus amendment from the previous time point. This should be a more accurate representation of OTUs responding to the amendment for each 

```{r, message=F, error=F, warning=F}
# RDS saved from 
resp_alf <- readRDS("../data/LFC_alf_OTUs.RDS") %>% dplyr::rename(label = OTU)
resp_comp <- readRDS("../data/LFC_comp_OTUs.RDS") %>% dplyr::rename(label = OTU)
resp_mix <- readRDS("../data/LFC_mix_OTUs.RDS") %>% dplyr::rename(label = OTU)

resp <- rbind(resp_alf[1], resp_comp[1], resp_mix[1])

dist_all_resp <- distinct(resp)

df <- resp_alf %>%
  full_join(resp_comp) %>%
  full_join(resp_mix)
```

```{r, message=F, error=F, warning=F}
inc_resp <- prune_taxa(c(df$label), inc) %>%
  filter_taxa(function(x) sum(x) >= 1, T)
```

```{r, message=F, error=F, warning=F}
inc_resp <- phy_tree(inc_resp)
inc_resp
```
```{r, message=F, error=F, warning=F}
df2 = dplyr::mutate(df, newlab = paste(label, Genus, sep='|'))
head(df2)
```
```{r, message=F, error=F, warning=F}
resp_p <- ggtree(inc_resp, branch.length = "none", aes(color=Phylum)) %<+% df2 +
  geom_tiplab(aes(label=newlab), size=3, align=TRUE, linesize=.5) +
  theme_tree() 
```
```{r, message=F, error=F, warning=F}
colnames(df2)
head(df2$label)

df3 <- df2 %>%
  select(label, Alfalfa_Early = Alfalfa_early_log2FoldChange, Alfalfa_Late = Alfalfa_late_log2FoldChange, Compost_Early = Compost_early_log2FoldChange, Compost_Late = Compost_late_log2FoldChange, Mix_Early = Mix_early_log2FoldChange, Mix_Late = Mix_late_log2FoldChange) %>%
  column_to_rownames(var = "label")

write.csv(df3, file ="../data/resp_table.csv")
final_resp <- resp_p %>% 
  gheatmap(df3, offset= 35, width= .8, font.size = 2.5, colnames=T, colnames_angle=-45, hjust=0, color = "white") +
  ggtitle("Log fold change of genera for each treatment response group. \nMin LFC = 4 when comparing to reference")

```
```{r, message=F, error=F, warning=F}

tree_plot <- final_resp + scale_fill_viridis(option = "plasma") +
  guides(fill=guide_legend(title="Log fold change"))

ggsave("../Figures/resp_tree_final.png", plot = tree_plot, device = "png", width = 18.5, height = 32, units = "cm", dpi = 350)
```

## Table of Otus    
All Otus with LFC greater than 4 in all treatment response groups
```{r, message=F, error=F, warning=F}
# Common responders in all amendments and response groups
new_DF <- df3[rowSums(is.na(df3)) == 0,]
kable(new_DF) %>%
  kable_styling(bootstrap_options = "striped") %>%
  scroll_box(width = "100%")
```

```{r}
p <- tree_plot  + 
  geom_hilight(node = 232, fill = "gray", alpha=0.4, extend = 36) + 
  geom_hilight(node = 201, fill = "gray", alpha=0.4, extend = 36) + 
  geom_hilight(node = 212, fill = "gray", alpha=0.4, extend = 36) + 
  geom_hilight(node = 183, fill = "gray", alpha=0.4, extend = 36) + 
  geom_hilight(node = 186, fill = "gray", alpha=0.4, extend = 36) + 
  geom_hilight(node = 193, fill = "gray", alpha=0.4, extend = 36) + 
  geom_hilight(node = 158, fill = "gray", alpha=0.4, extend = 36) + 
  geom_hilight(node = 132, fill = "gray", alpha=0.4, extend = 36) + 
  geom_hilight(node = 124, fill = "gray", alpha=0.4, extend = 36) + 
  geom_hilight(node = 126, fill = "gray", alpha=0.4, extend = 36) +
  geom_label2(aes(subset = node %in% c(232,201,212,183,186,193,158,132,124,126), label=node), size=2, color="black", fill="gray", alpha = 1)

z <- tree_plot +
  geom_label2(aes(subset = !isTip, label=node), size=2, color="darkblue", alpha=1) 
p
ggsave("../Figures/resp_tree_wnodes.png", plot = p, device = "png", width = 18.5, height = 32, units = "cm", dpi = 350)
ggsave("../Figures/resp_tree_allnodes.png", plot = z, device = "png", width = 18.5, height = 32, units = "cm", dpi = 350)
```



# Responders orginating from amendment? 

Aliens??

```{r}
df3 <- read_csv("../data/resp_table.csv")
df3
treeOTUs <- df3$X1
alfALIENs <- readRDS("../data/alf.aliens.rds")
compALIENs <- readRDS("../data/comp.aliens.rds")
mixALIENs <- readRDS("../data/mix.aliens.rds")

venn(list(treeOTUs, alfALIENs, compALIENs, mixALIENs), snames = c("Responder", "Alf Aliens", "Compost Aliens", "Mix Aliens"))

venn(list(treeOTUs, alfALIENs), snames = c("Responder", "Alf Aliens"))
a <- venn(list(treeOTUs, alfALIENs), snames = c("Responder", "Alf Aliens"))


venn(list(treeOTUs, compALIENs), snames = c("Responder", "Compost Aliens"))
c <- venn(list(treeOTUs, compALIENs), snames = c("Responder", "Compost Aliens"))
# Alfalfa
print("Alfalfa Aliens")
attr(a, "intersections")$`Responder:Alf Aliens`

# Compost
print("Compost Aliens")
attr(c, "intersections")$`Responder:Compost Aliens`
```
    
We have 3 OTUs from compost and alfalfa that are detected as responders! Let's label them on the tree plot

```{r}
# Make these green, orgianlly from alfalfa
p$data %>%
  filter(label %in% c("Otu01028", "Otu01189",  "Otu03173")) %>%
  select(node)
# Make these brown, orgianlly from compost
p$data %>%
  filter(label %in% c("Otu00887", "Otu00985", "Otu01490")) %>%
  select(node)

t <- p + 
  geom_hilight(node = 4, fill = "green", alpha=0.4, extend = 36) + 
  geom_hilight(node = 24, fill = "green", alpha=0.4, extend = 36) + 
  geom_hilight(node = 32, fill = "green", alpha=0.4, extend = 36) + 
  geom_hilight(node = 62, fill = "brown", alpha=0.4, extend = 36) + 
  geom_hilight(node = 64, fill = "brown", alpha=0.4, extend = 36) + 
  geom_hilight(node = 65, fill = "brown", alpha=0.4, extend = 36)
ggsave("../Figures/resp_tree_walienshilighted.png", plot = t, device = "png", width = 18.5, height = 32, units = "cm", dpi = 350)
t
```

