---
title: "Plotting of relative abundance of responding phyla to amendment in the Incubation Project"
output:
  html_document:
    df_print: paged
    fig_width: 8.5
    fig_height: 11
    fig_caption: true
    toc: true
    toc_float: true
    number_sections: false
    code_folding: hide
---
```{r setup}
knitr::opts_knit$set(root.dir = normalizePath("."))
```

```{r BiocManager, cache=TRUE}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install(c("phyloseq", "DESeq2"))
```

```{r Libraries, echo = F, message = F, warning = F}
# Load libraries
library(tidyverse)
library(phyloseq)
library(DESeq2)
library(naniar)
library(ggrepel)
library(viridis)
library(knitr)
library(kableExtra)
library(gridExtra)
```

```{r Phyloseq Object}
inc.physeq <- readRDS("../data/not.rare.nounclass")
source("../code/functions.R")
```

```{r Alf OTUs}
early_alf_OTUS <- readRDS("../data/LFC>4_early_alf_OTUs.RDS")
late_alf_OTUS <- readRDS("../data/LFC>4_late_alf_OTUs.RDS")
all_alf <- readRDS("../data/LFC_alf_OTUs.RDS")
```

```{r Comp OTUs}
early_comp_OTUS <- readRDS("../data/LFC>4_early_comp_OTUs.RDS")
late_comp_OTUS <- readRDS("../data/LFC>4_late_comp_OTUs.RDS")
all_comp <- readRDS("../data/LFC_comp_OTUs.RDS")
```

```{r Mix OTUs}
early_mix_OTUS <- readRDS("../data/LFC>4_early_mix_OTUs.RDS")
late_mix_OTUS <- readRDS("../data/LFC>4_late_mix_OTUs.RDS")
all_mix <- readRDS("../data/LFC_mix_OTUs.RDS")
```

```{r consistent coloring}
# We have many more colors need, it would be easiest to have one that can cover all levels of phyla
# list of phyla
# We also need to keep this consistent accross treatments, so we need the phyla from each treatment not just alf
list <- as.character(unique(all_alf$Phylum)) %>%
  c(as.character(unique(all_comp$Phylum))) %>%
  c(as.character(unique(all_mix$Phylum)))

phylalist <- sample(unique(list))  # note the sample(), this randomizes the list, if you leave it you may find colors to similar that are close to each other. 

# this package will generate a pallette based on number and desired colors
library(colorspace)
colors27 <- sequential_hcl(n_distinct(phylalist), palette = "viridis")

# I found this on stackoverflow
# It's a function that will assign a color to a specific phylum
scale_fill_phyla_consistent <- function(...){
    ggplot2:::manual_scale(
        'fill', 
        values = setNames(colors27, phylalist), 
        ...
    )
}
```

# Alfalfa Area

Do the responding OTUs always represent a significant relative abundance?

Let's visualize the relative abundance of the significantly responding genera in each treatment. We will use a stacked area chart.

```{r psmelt alfalfa, warning=FALSE}
# We want to visualise the alfalfa treatment responding OTUs from early and late to reference. 
# Use two plots faceted by treatment
alf <- subset_samples(inc.physeq, treatment %in% c("Alfalfa") & day %in% c("0", "7", "14", "21", "35", "49", "97")) 

ref <- subset_samples(inc.physeq, treatment %in% c("Reference") & day %in% c("0", "7", "14", "21", "35", "49", "97")) 

# We have multiple replicates for each combination of treatment and day, here we merge the samples from each combination
# merge messes with the columns, we can fix with sample_data()
rela.phy <- merge_samples(alf, "day") 
rela.phy.data <- data.frame(sample_data(rela.phy))
sample_data(rela.phy)$day = sample_names(rela.phy)
sample_data(rela.phy)$treatment = "Alfalfa"

ref.phy <- merge_samples(ref, "day")
ref.phy.data <- data.frame(sample_data(ref.phy))
sample_data(ref.phy)$day = sample_names(ref.phy)
sample_data(ref.phy)$treatment = "Reference"

# melt each into a data structure
meltalf <- rela.phy %>%
  transform_sample_counts(function(x) x / sum(x)) %>%
  psmelt() %>%
  select(OTU, Abundance, treatment, day, Phylum)

meltref <- ref.phy %>%
  transform_sample_counts(function(x) x / sum(x)) %>%
  psmelt() %>%
  select(OTU, Abundance, treatment, day, Phylum)
```

```{r rbind alfalfa, warning=FALSE} 
# We need a data frame with the relative abundance of the responding Phyla
data <- rbind(meltalf, meltref) %>%
  mutate(Responder = ifelse(OTU %in% early_alf_OTUS$OTU, "Responding", "Not_responding")) %>%
  filter(Responder == "Responding") %>%
  group_by(treatment, day, Phylum) %>%
  mutate(phyla_sum = sum(Abundance)) %>%
  select(treatment, day, Phylum, phyla_sum) %>%
  distinct()

datalate <- rbind(meltalf, meltref) %>%
  mutate(Responder = ifelse(OTU %in% late_alf_OTUS$OTU, "Responding", "Not_responding")) %>%
  filter(Responder == "Responding") %>%
  group_by(treatment, day, Phylum) %>%
  mutate(phyla_sum = sum(Abundance)) %>%
  select(treatment, day, Phylum, phyla_sum) %>%
  distinct()
```

```{r Sort alfalfa and plot}
# We can sort the data based on abundance at day 7, this makes a cleaner graph
myLevels <- data %>%
  filter(day==7 & treatment == "Alfalfa")%>%
  arrange(desc(phyla_sum))

data$Phylum <- factor(data$Phylum, levels = myLevels$Phylum) 

pp <- data %>%
  ggplot(aes(x= as.numeric(day), y=phyla_sum)) + 
  geom_area(aes(fill = Phylum), size=.25, colour="black") + 
  facet_grid(treatment ~ .) 
  
#pp

ppp <- pp + 
  scale_x_continuous(breaks = c(0, 7, 14, 21, 35, 49, 97)) +
  xlab("Day") +
  ylab("Relative abundance \n of responders (LFC > 4)") +
  ggtitle("Early alfalfa responders") 
#ppp
```

```{r late alfalfa and plot}
# Sort based on abundance at day 35, determine order
myLevels2 <- datalate %>%
  filter(day==35 & treatment == "Alfalfa") %>%
  arrange(desc(phyla_sum))

# Set order of Phyla based on abundance at day 35
datalate$Phylum <- factor(datalate$Phylum, levels = myLevels2$Phylum) 
  
plot2 <- datalate %>%
  ggplot(aes(x= as.numeric(day), y=phyla_sum)) + 
  geom_area(aes(fill = Phylum), size=.25, colour="black") + 
  facet_grid(treatment ~ .) 

#plot2

# Add titles and scale x-axis
pplot2 <- plot2 + 
  scale_x_continuous(breaks = c(0, 7, 14, 21, 35, 49, 97)) +
  xlab("Day") +
  ylab("Relative abundance \n of responders (LFC > 4)") +
  ggtitle("Late alfalfa responders") 

#pplot2
```

Note on coloring, it's annoying how phyla with have different colors between plots, let's see if we can fix that
```{r plot alf with consistent colors}
ppp + scale_fill_phyla_consistent()
pplot2 + scale_fill_phyla_consistent()
# Save 
png("../Figures/alfalfa_early_area_responders.png",height=8,width=10,units='in',res=300)
ppp + scale_fill_phyla_consistent()
dev.off()
png("../Figures/alfalfa_late_area_responders.png",height=8,width=10,units='in',res=300)
pplot2 + scale_fill_phyla_consistent()
dev.off()
```

# Compost Area

```{r psmelt Compost}
comp <- subset_samples(inc.physeq, treatment %in% c("Compost") & day %in% c("0", "7", "14", "21", "35", "49", "97")) 

rela.phy <- merge_samples(comp, "day") 
rela.phy.data <- data.frame(sample_data(rela.phy))
sample_data(rela.phy)$day = sample_names(rela.phy)
sample_data(rela.phy)$treatment = "Compost"

# melt into a data structure
meltcomp <- rela.phy %>%
  transform_sample_counts(function(x) x / sum(x)) %>%
  psmelt() %>%
  select(OTU, Abundance, treatment, day, Phylum)
```

```{r rbind Compost} 
# We need a data frame with the relative abundance of the responding Phyla
data <- rbind(meltcomp, meltref) %>%
  mutate(Responder = ifelse(OTU %in% early_comp_OTUS$OTU, "Responding", "Not_responding")) %>%
  filter(Responder == "Responding") %>%
  group_by(treatment, day, Phylum) %>%
  mutate(phyla_sum = sum(Abundance)) %>%
  select(treatment, day, Phylum, phyla_sum) %>%
  distinct()

datalate <- rbind(meltcomp, meltref) %>%
  mutate(Responder = ifelse(OTU %in% late_comp_OTUS$OTU, "Responding", "Not_responding")) %>%
  filter(Responder == "Responding") %>%
  group_by(treatment, day, Phylum) %>%
  mutate(phyla_sum = sum(Abundance)) %>%
  select(treatment, day, Phylum, phyla_sum) %>%
  distinct()
```

```{r Reorder compost and plot}
# We can sort the data based on abundance at day 7, this makes a cleaner graph
myLevels <- data %>%
  filter(day==7 & treatment == "Compost")%>%
  arrange(desc(phyla_sum))

data$Phylum <- factor(data$Phylum, levels = myLevels$Phylum) 

pp <- data %>%
  ggplot(aes(x= as.numeric(day), y=phyla_sum)) + 
  geom_area(aes(fill = Phylum), size=.25, colour="black") + 
  facet_grid(treatment ~ .) 
  
#pp

ppp <- pp + 
  scale_x_continuous(breaks = c(0, 7, 14, 21, 35, 49, 97)) +
  xlab("Day") +
  ylab("Relative abundance \n of responders (LFC > 4)") +
  ggtitle("Early Compost responders") 
#ppp
```

```{r late compost order and sort}
# Sort based on abundance at day 35, determine order
myLevels2 <- datalate %>%
  filter(day==35 & treatment == "Compost") %>%
  arrange(desc(phyla_sum))

# Set order of Phyla based on abundance at day 35
datalate$Phylum <- factor(datalate$Phylum, levels = myLevels2$Phylum) 
  
plot2 <- datalate %>%
  ggplot(aes(x= as.numeric(day), y=phyla_sum)) + 
  geom_area(aes(fill = Phylum), size=.25, colour="black") + 
  facet_grid(treatment ~ .) 

#plot2

# Add titles and scale x-axis
pplot2 <- plot2 + 
  scale_x_continuous(breaks = c(0, 7, 14, 21, 35, 49, 97)) +
  xlab("Day") +
  ylab("Relative abundance \n of responders (LFC > 4)") +
  ggtitle("Late Compost responders") 

#pplot2
```

Note on coloring, it's annoying how phyla with have different colors between plots, let's see if we can fix that
```{r Coloring compost plot}
ppp + scale_fill_phyla_consistent()
pplot2 + scale_fill_phyla_consistent()

# Save 
png("../Figures/compost_early_area_responders.png",height=8,width=10,units='in',res=300)
ppp + scale_fill_phyla_consistent()
dev.off()
png("../Figures/compost_late_area_responders.png",height=8,width=10,units='in',res=300)
pplot2 + scale_fill_phyla_consistent()
dev.off()
```

# Mix Area

```{r psmelt Mix, warning=FALSE}
mix <- subset_samples(inc.physeq, treatment %in% c("Mix") & day %in% c("0", "7", "14", "21", "35", "49", "97")) 

rela.phy <- merge_samples(mix, "day") 
rela.phy.data <- data.frame(sample_data(rela.phy))
sample_data(rela.phy)$day = sample_names(rela.phy)
sample_data(rela.phy)$treatment = "mix"

# melt into a data structure
meltmix <- rela.phy %>%
  transform_sample_counts(function(x) x / sum(x)) %>%
  psmelt() %>%
  select(OTU, Abundance, treatment, day, Phylum)
```

```{r rbind Mix} 
# We need a data frame with the relative abundance of the responding Phyla
data <- rbind(meltmix, meltref) %>%
  mutate(Responder = ifelse(OTU %in% early_mix_OTUS$OTU, "Responding", "Not_responding")) %>%
  filter(Responder == "Responding") %>%
  group_by(treatment, day, Phylum) %>%
  mutate(phyla_sum = sum(Abundance)) %>%
  select(treatment, day, Phylum, phyla_sum) %>%
  distinct()

datalate <- rbind(meltmix, meltref) %>%
  mutate(Responder = ifelse(OTU %in% late_mix_OTUS$OTU, "Responding", "Not_responding")) %>%
  filter(Responder == "Responding") %>%
  group_by(treatment, day, Phylum) %>%
  mutate(phyla_sum = sum(Abundance)) %>%
  select(treatment, day, Phylum, phyla_sum) %>%
  distinct()
```

```{r Reorder Mix and Plot}
# We can sort the data based on abundance at day 7, this makes a cleaner graph
myLevels <- data %>%
  filter(day==7 & treatment == "mix")%>%
  arrange(desc(phyla_sum))

data$Phylum <- factor(data$Phylum, levels = myLevels$Phylum) 

pp <- data %>%
  ggplot(aes(x= as.numeric(day), y=phyla_sum)) + 
  geom_area(aes(fill = Phylum), size=.25, colour="black") + 
  facet_grid(treatment ~ .) 
  
#pp

ppp <- pp + 
  scale_x_continuous(breaks = c(0, 7, 14, 21, 35, 49, 97)) +
  xlab("Day") +
  ylab("Relative abundance \n of responders (LFC > 4)") +
  ggtitle("Early mix responders") 
#ppp
```

```{r Late mix reorder and plot}
# Sort based on abundance at day 35, determine order
myLevels2 <- datalate %>%
  filter(day==35 & treatment == "mix") %>%
  arrange(desc(phyla_sum))

# Set order of Phyla based on abundance at day 35
datalate$Phylum <- factor(datalate$Phylum, levels = myLevels2$Phylum) 
  
plot2 <- datalate %>%
  ggplot(aes(x= as.numeric(day), y=phyla_sum)) + 
  geom_area(aes(fill = Phylum), size=.25, colour="black") + 
  facet_grid(treatment ~ .) 

#plot2

# Add titles and scale x-axis
pplot2 <- plot2 + 
  scale_x_continuous(breaks = c(0, 7, 14, 21, 35, 49, 97)) +
  xlab("Day") +
  ylab("Relative abundance \n of responders (LFC > 4)") +
  ggtitle("Late mix responders") 

#pplot2
```

Note on coloring, it's annoying how phyla with have different colors between plots, let's see if we can fix that
```{r Color mix plot}
ppp + scale_fill_phyla_consistent()
pplot2 + scale_fill_phyla_consistent()

# Save 
png("../Figures/mix_early_area_responders.png",height=8,width=10,units='in',res=300)
ppp + scale_fill_phyla_consistent()
dev.off()
png("../Figures/mix_late_area_responders.png",height=8,width=10,units='in',res=300)
pplot2 + scale_fill_phyla_consistent()
dev.off()
```

